oscar_figures

### 2021-03-14 oscar_figures.Rmd
Add iCLIP peak plots

@@

## iCLIP

>>>>

TODO: add scripts which deduplicate UMIs for each.

@@

## iCLIP of minigenes

Read in the processed data (following alignment with Bowtie2, and filtering for high-confidence alignments), and generate
a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy

TODO: add script which does this filtering

```{r message=FALSE, warning=FALSE}
library(patchwork)

exon_snp <- 556 
intron_snp <- 1106
intron_start <- 141 
intron_end <- 1429
acceptor1 <- 394
acceptor2 <- 444
donor <- 572
mg_l <- 1624 
cpoor_start <- 490
cpoor_end <- 1325
risk_colour <- "#00BFC4"
healthy_colour <- "#F8766D"

filtered_df <- read_csv(paste0(here(), "/data/iCLIP/bowtie2_aligned_2_filtered_df.csv"))

rolling_window <- 20

df2 <- filtered_df %>%
  ungroup() %>%
  group_by(condition, pos) %>%
  mutate(mean_counts = sum(normalised_counts)/2) %>%
  select(pos, condition, mean_counts) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from ="mean_counts")

df2[is.na(df2)] <- 0

df2 <- df2 %>% 
  mutate(diff = R-H) %>%
  mutate(mean_4 = (H+R)/2) %>%
  arrange(pos)

top <- ggplot(df2, aes(x=pos, y=diff)) +
  geom_area(aes(x=pos, y = mean_4), colour="grey50", fill = "grey50", alpha=0.2) +
  geom_vline(xintercept = exon_snp, linetype="dashed") +
  geom_vline(xintercept = intron_snp, linetype="dashed") +
  xlim(1,1600) +
  #ylim(-1,5) +
  theme_minimal() +
  ylab("XLs per 1000") +
  xlab("") +
  geom_vline(xintercept = cpoor_start) +
  geom_vline(xintercept = cpoor_end)

bottom <- ggplot(df2, aes(x=pos, y=zoo::rollmean(diff, k = rolling_window, na.pad=T))) +
  geom_area(colour = "grey50", fill="grey50", alpha = 1) +
  geom_vline(xintercept =  exon_snp, linetype="dashed") +
  geom_vline(xintercept =  intron_snp, linetype="dashed") +
  xlim(1,1600) +
  ylim(-0.5,0.75) +
  theme_minimal() +
  ylab("Change in XL density") +
  xlab("Distance from start of minigene") +
  geom_vline(xintercept =  intron_start) +
  geom_vline(xintercept =  intron_end) +
  geom_vline(xintercept =  acceptor1) +
  geom_vline(xintercept =  acceptor2) +
  geom_vline(xintercept =  donor)

combined <- top/bottom
combined
```

Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.

```{r message=FALSE, warning=FALSE}

find_ratio_bound <- function(e_x, sem_x, e_y, sem_y, quantile_value, randoms = 10000){
  # monte carlo confidence interval estimate for a ratio. Assumes the two variables
  # x and y are normally distributed. Finds confidence interval for (x-y)/y

  # randomly generate possible global means for x and y

  xs = rnorm(randoms, mean = e_x, sd=sem_x)  # risk
  ys = rnorm(randoms, mean = e_y, sd=sem_y)  # healthy

  ratios = (xs-ys)/ys

  return(quantile(ratios, quantile_value))
}

## Plot big changes

confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.

peak_threshold <- 5 # how many times above the mean to be a "peak"
region_length <- 50

peak_t <- peak_threshold*1000/mg_l

peak_df <- filtered_df %>%
  ungroup() %>%
  group_by(pos) %>%
  mutate(max_at_pos = max(normalised_counts)) %>%
  ungroup() %>%
  filter(max_at_pos >= peak_t) %>%
  group_by(condition, pos) %>%
  mutate(peak_mean = mean(normalised_counts), 
         peak_sd = sd(normalised_counts)) %>%
  ungroup() %>%
  select(condition, pos, peak_mean, peak_sd) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from = c("peak_mean", "peak_sd")) %>%
  mutate(frac_increase = (peak_mean_R-peak_mean_H)/peak_mean_H) %>%
  mutate(peak_rank = rank(frac_increase)) %>%
  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "gold", 
                         ifelse(abs(pos-intron_snp)<region_length, "purple", "grey50")))

peak_df$upper_bound <- mapply(find_ratio_bound, 
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1+confidence_interval))

peak_df$lower_bound <- mapply(find_ratio_bound, 
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1-confidence_interval))

ggplot(peak_df, aes(x = peak_rank, y = 100*frac_increase, fill=colour)) +
  geom_bar(stat="identity") +
  scale_fill_identity() +
  theme_minimal() +
  ggpubr::theme_pubr()+
  ylab("% change in 2x Risk") +
  xlab("Rank") +
  geom_errorbar(aes(ymin = 100*lower_bound, ymax = 100*upper_bound, x = peak_rank), 
                width = 0.3)
